---
title: "Question6ToothGrowth"
author: "Justice Ruwona"
date: "2023-03-08"
output: 
  word_document: default
  html_document:
    keep_md: yes
---

## Load the datset
```{r}
data(ToothGrowth)
str(ToothGrowth)
```

## Using Subset method

### Subset ToothGrowth to include rows such that supp is equal to VC:

```{r}
VC_data <- subset(ToothGrowth, supp == "VC")
str(VC_data)
```
### Subset ToothGrowth to include rows such that supp is equal to VC and dose is equal to 0.5:
```{r}
VC_dose_data <- subset(ToothGrowth, supp == "VC" & dose == 0.5)
str(VC_dose_data)
```


### Subset ToothGrowth to include the values of len such that supp is equal to VC and dose is equal to 0.5:

```{r}
VC_dose_len <- subset(ToothGrowth, supp == "VC" & dose == 0.5, select = len)
str(VC_dose_len)
```

## Using Indexing method
### Subset ToothGrowth to include rows such that supp is equal to VC:
```{r}
toothgrowth_vc <- ToothGrowth[ToothGrowth$supp == "VC", ]
str(toothgrowth_vc)
```
### Subset ToothGrowth to include rows such that supp is equal to VC and dose is equal to 0.5:
```{r}
toothgrowth_vc_dose <- ToothGrowth[ToothGrowth$supp == "VC" & ToothGrowth$dose == 0.5, ]
str(toothgrowth_vc_dose)
```
### Subset ToothGrowth to include the values of len such that supp is equal to VC and dose is equal to 0.5:
```{r}
toothgrowth_vc_dose_len <- ToothGrowth$len[ToothGrowth$supp == "VC" & ToothGrowth$dose == 0.5]
str(toothgrowth_vc_dose_len)
```




