---
title: "Data Visualization Assignment"
author: "Justice Ruwona"
date: "2023-02-26"
output: 
  html_document:
    keep_md: yes
  word_document: default
---

## Load relevant libraries for visualizations
```{r, include=TRUE}
library(knitr)
library(ggplot2)
library(markdown)
```

# #  get working directory to confirm location of file
```{r}
getwd()
```
# 2. Make a boxplot using ggplot with DON as the y variable, treatment as the x variable, and color mapped to the wheat cultivar. Show the code you use to load the libraries you need read in the data and make the plot. Change the y label to “DON (ppm)” and make the x label blank

# #Read in data,load ggplot library and plot boxplot
```{r}
MycotoxinData <- read.csv("MycotoxinData.csv", na.strings = "na", header = TRUE)

# check if data read in correctly
 str(MycotoxinData)
 
#Plot boxplot
ggplot(MycotoxinData, aes(x = Treatment, y = DON, color = Cultivar))+
    geom_boxplot() +
    ylab("DON (ppm)") + # y label DON
    xlab("") + #x label blank
    scale_fill_discrete(name = "Wheat Cultivar")
    
```



#3. Now convert this same data into a bar chart with standard-error error bars

## Make bar chart using the Mycotoxin data with standard error bars
```{r}
ggplot(MycotoxinData, aes(x = Treatment, y = DON, color = Cultivar, fill = Cultivar))+
    stat_summary(fun=mean,geom="bar", position = "dodge") +
    stat_summary(fun.data = mean_se, geom = "errorbar", position = "dodge") + #calculate means and standard errors for each group
    ylab("DON (ppm)") +
    xlab("")
```
# 4 . Add points to the foreground of the boxplot and bar chart you made in questions 2 and 3 that show the distribution of points over the boxplots or bars. Set the shape = 21 and outline color black (hint: use jitter_dodge). 

# # Boxplot with points to the foreground
```{r}
ggplot(MycotoxinData, aes(x = Treatment, y = DON, color = Cultivar))+
    geom_boxplot(color = "black") + # change outline to black
    geom_jitter(aes(color = Cultivar), shape = 21, size = 3, position = position_jitterdodge(jitter.width = 0.2, jitter.height = 0)) +
    ylab("DON (ppm)") +
    xlab("") +
    scale_fill_discrete(name = "Cultivar")
    ```
# # Bar chart with points to the foreground and outline color black
```{r}
library(ggplot2)
ggplot(MycotoxinData, aes(x = Treatment, y = DON, color = Cultivar, fill = Cultivar))+
    stat_summary(fun=mean,geom="bar", position = "dodge", color = "black") + # change outline to black
    stat_summary(fun.data = mean_se, geom = "errorbar", position = "dodge") +
    geom_jitter(aes(color = Cultivar), shape = 21, size = 3, position = position_jitterdodge(jitter.width = 0.2, jitter.height = 0)) +
    ylab("DON (ppm)") +
    xlab("")
```
# 5. Change the color and fill of the points and boxplots to match some of the colors in the following colorblind pallet. cbbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")


```{r}
# The palette with black:
cbbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

#boxplot
ggplot(MycotoxinData, aes(x = Treatment, y = DON, color = Cultivar))+
    geom_boxplot(color = "black") +
    geom_jitter(aes(color = Cultivar), shape = 21, size = 3, position = position_jitterdodge(jitter.width = 0.2, jitter.height = 0)) +
    ylab("DON (ppm)") +
    xlab("") +
    scale_fill_discrete(name = "Cultivar")+
    scale_fill_manual(values = cbbPalette) +
    scale_color_manual(values = cbbPalette) 

#barchat
ggplot(MycotoxinData, aes(x = Treatment, y = DON, color = Cultivar, fill = Cultivar))+
    stat_summary(fun=mean,geom="bar", position = "dodge", color = "black") +
    stat_summary(fun.data = mean_se, geom = "errorbar", position = "dodge") +
    geom_jitter(aes(color = Cultivar), shape = 21, size = 3, position = position_jitterdodge(jitter.width = 0.2, jitter.height = 0)) +
    ylab("DON (ppm)") +
    xlab("")+
    scale_fill_manual(values = cbbPalette) +
    scale_color_manual(values = cbbPalette) 
```
# 6. Add a facet to the plots based on cultivar
```{r}
# The palette with black:
cbbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

#boxplot
ggplot(MycotoxinData, aes(x = Treatment, y = DON, color = Cultivar))+
    geom_boxplot(color = "black") +
    geom_jitter(aes(color = Cultivar), shape = 21, size = 3, position = position_jitterdodge(jitter.width = 0.2, jitter.height = 0)) +
    ylab("DON (ppm)") +
    xlab("") +
    scale_fill_discrete(name = "Cultivar")+
    scale_fill_manual(values = cbbPalette) +
    scale_color_manual(values = cbbPalette) +
    facet_wrap(~Cultivar)

#barchart
ggplot(MycotoxinData, aes(x = Treatment, y = DON, color = Cultivar, fill = Cultivar))+
    stat_summary(fun=mean,geom="bar", position = "dodge", color = "black") +
    stat_summary(fun.data = mean_se, geom = "errorbar", position = "dodge") +
    geom_jitter(aes(color = Cultivar), shape = 21, size = 3, position = position_jitterdodge(jitter.width = 0.2, jitter.height = 0)) +
    ylab("DON (ppm)") +
    xlab("")+
    scale_fill_manual(values = cbbPalette) +
    scale_color_manual(values = cbbPalette) +
    facet_wrap(~Cultivar)
```
# 7. Explore the other themes and change to a theme of interest. https://ggplot2.tidyverse.org/reference/ggtheme.html

# change boxplot to theme_dark and barchart to theme_linedraw
```{r}
# The palette with black:
cbbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

#boxplot change to 
ggplot(MycotoxinData, aes(x = Treatment, y = DON, color = Cultivar))+
    geom_boxplot(color = "black") +
    geom_jitter(aes(color = Cultivar), shape = 21, size = 3, position = position_jitterdodge(jitter.width = 0.2, jitter.height = 0)) +
    ylab("DON (ppm)") +
    xlab("") +
    scale_fill_discrete(name = "Cultivar")+
    scale_fill_manual(values = cbbPalette) +
    scale_color_manual(values = cbbPalette) +
    facet_wrap(~Cultivar) +
    theme_dark()
```
```{r}
#barchart
ggplot(MycotoxinData, aes(x = Treatment, y = DON, color = Cultivar, fill = Cultivar))+
    stat_summary(fun=mean,geom="bar", position = "dodge", color = "black") +
    stat_summary(fun.data = mean_se, geom = "errorbar", position = "dodge") +
    geom_jitter(aes(color = Cultivar), shape = 21, size = 3, position = position_jitterdodge(jitter.width = 0.2, jitter.height = 0)) +
    ylab("DON (ppm)") +
    xlab("")+
    scale_fill_manual(values = cbbPalette) +
    scale_color_manual(values = cbbPalette) +
    facet_wrap(~Cultivar) +
    theme_linedraw()
```



# 8. Add transparency to the points so you can still see the boxplot or bar in the background. 
```{r}

#boxplot seen in the background
ggplot(MycotoxinData, aes(x = Treatment, y = DON, color = Cultivar))+
    geom_boxplot(color = "black") +
    geom_jitter(aes(color = Cultivar), shape = 21, size = 3, position = position_jitterdodge(jitter.width = 0.2, jitter.height = 0)) +
    ylab("DON (ppm)") +
    xlab("") +
    scale_fill_discrete(name = "Cultivar")+
    theme_bw()

```
```{r}
#barchart with transparecy
library(ggplot2)
ggplot(MycotoxinData, aes(x = Treatment, y = DON, color = Cultivar, fill = Cultivar))+
    stat_summary(fun=mean,geom="bar", position = "dodge", color = "black") + # change outline to black
    stat_summary(fun.data = mean_se, geom = "errorbar", position = "dodge") +
    geom_jitter(aes(color = Cultivar), shape = 21, size = 3, alpha =0.4, position = position_jitterdodge(dodge.width = 0.1, jitter.width = 0.25)) + # using position jitter to make transparency
    ylab("DON (ppm)") +
    xlab("") +
    theme_bw()
```
# 10. Exploring other ways to visaulize the data

# using dotplot instead of geom_boxplot

# using statcount instead of barchart
```{r}
#using geom_point

ggplot(MycotoxinData, aes(x = Treatment, y = DON, color = Cultivar))+
    geom_point() +
    ylab("DON (ppm)") +
    xlab("") +
    scale_fill_discrete(name = "Cultivar")+
    theme_bw()
```
```{r}
#using stat_count
ggplot(MycotoxinData, aes(x = Treatment, fill = Cultivar)) +
  stat_count(alpha = 0.5, position = "identity") +
  ylab("DON (ppm)") +
  xlab("") +
  scale_fill_discrete(name = "Cultivar") +
  theme_bw()
```

