---
title: "Data Visualization Assignment"
author: "Justice Ruwona"
date: "2023-02-26"
output: 
  html_document:
    keep_md: yes
  word_document: default
---
```{r, include=TRUE}
library(knitr)
library(ggplot2)
library(markdown)
```

# #  get working directory to confirm location of file
```{r}
getwd()
```
# 2. Make a boxplot using ggplot with DON as the y variable, treatment as the x variable, and color mapped to the wheat cultivar. Show the code you use to load the libraries you need read in the data and make the plot. Change the y label to “DON (ppm)” and make the x label blank

# #Read in data,load ggplot library and plot boxplot
```{r}
library(ggplot2)

MycotoxinData <- read.csv("MycotoxinData.csv", na.strings = "na", header = TRUE)
sum(is.na(MycotoxinData))

#make plot
ggplot(MycotoxinData, aes(x = Treatment, y = DON, color = Cultivar))+
    geom_boxplot() +
    ylab("DON (ppm)") +
    xlab("") +
    scale_fill_discrete(name = "Wheat Cultivar")
    
```



#3. Now convert this same data into a bar chart with standard-error error bars

## Make bar chart using the Mycotoxin data with standard error bars
```{r}
library(ggplot2)
#calculate means and standard errors for each group
 ggplot(MycotoxinData, aes(x = Treatment, y = DON, color = Cultivar, fill = Cultivar))+
    stat_summary(fun=mean,geom="bar", position = "dodge") +
    stat_summary(fun.data = mean_se, geom = "errorbar", position = "dodge") +
    ylab("DON (ppm)") +
    xlab("")
```
```{r}
# Add points to the boxplot
ggplot(MycotoxinData, aes(x = Treatment, y = DON, color = Cultivar))+
    geom_boxplot() +
    geom_jitter(aes(shape=Cultivar), position=position_jitterdodge(jitter.width=0.2, dodge.width=0.9), size=2, stroke=0.5) +
    ylab("DON (ppm)") +
    xlab("") +
    scale_fill_discrete(name = "Cultivar")

```
#add points to the bar chart using the Mycotoxin data with standard error bars
```{r}
library(ggplot2)
#calculate means and standard errors for each group
 ggplot(MycotoxinData, aes(x = Treatment, y = DON, color = Cultivar, fill = Cultivar))+
    stat_summary(fun=mean,geom="bar", position = "dodge") +
    stat_summary(fun.data = mean_se, geom = "errorbar", position = "dodge") +
    geom_jitter(aes(shape=Cultivar), position=position_jitterdodge(jitter.width=0.2, dodge.width=0.9), size=2, stroke=0.5) +
    ylab("DON (ppm)") +
    xlab("")
```
#Change the color and fill 
```{r}
```
